\section{Introduction}

\subsection{Background}

    \subsubsection{Four Campuses One Reality}
    % Something about the vision and idea behind this initiative
    It's about connection Gløs, Draggis, Gjøvik og Ålesund through VR.

    \subsubsection{IVR-Connection}
    % About this program and how it fits in with the vision and idea from 4C1R
    Program made by Niclas to test cooperation in VR.

    \subsubsection{Concept}
    % Something about our addition to the vision and idea of 4C1R.. mebe?
    We want to build on Niclas' concept and add Hololens into the mix.
    
    \subsubsection{State of the Art}
    % What is the latest tech and research

\subsection{Technology}
% What technology are we using. Mebe? Subject to change

    \subsubsection{Universal Windows Platform}
    % About UWP. What is it and why we use it
    Universal Windows Platform (UWP) is a platform-homogeneous application architecture created by Microsoft and first introduced in Windows 10. The purpose of this software platform is to help develop universal apps that run on Windows 10, Windows 10 Mobile, Xbox One and Hololens without the need to be re-written for each. It supports Windows app development using C++, C#, VB.NET, and XAML. The API is implemented in C++, and supported in C++, VB.NET, C#, F# and JavaScript.[1] Designed as an extension to the Windows Runtime platform first introduced in Windows Server 2012 and Windows 8, UWP allows developers to create apps that will potentially run on multiple types of devices. %Wikipedia. Probably cite.

    \subsubsection{Mixed Reality Toolkit}
    % What is it and why we use it
    The Mixed Reality Toolkit is a collection of scripts and components intended to accelerate development of applications targeting Microsoft HoloLens and Windows Mixed Reality headsets. The project is aimed at reducing barriers to entry to create mixed reality applications and contribute back to the community as we all grow. %Copy pasta from github readme

    \subsubsection{Unity}
    % What is it and why we use it
    Unity Technologies offers a platform for creating beautiful and engaging 2D, 3D, VR, and AR games and apps. A powerful graphics engine and full-featured editor enable you to realize your creative vision fast, and deliver your content to virtually any media or device. You can easily connect to your audiences on PCs, consoles, the web, mobile devices, home entertainment systems, embedded systems, or head-mounted displays.

    More than an engine, Unity helps you achieve ongoing success. It offers everything you need to develop quality content, boost your productivity, and connect with your audience. Tools and resources include the Unity Asset Store, Unity Cloud Build, Unity Analytics, Unity Ads, Unity Everyplay, and Unity Certification. Unity Technologies serves millions of registered developers including large publishers, indie studios, students and hobbyists around the globe.
    
    The leading global game industry software
    Unity plays an important part in a booming global games market. More games are made with Unity than with any other game technology. More players play games made with Unity, and more developers rely on our tools and services to drive their business.
    
    As you can see below, millions of people are creating with Unity every day. This has resulted in global gamers downloading Unity-made games to nearly 2 billion unique mobile devices for Q1 2016 alone.
    
    Unity is also at the forefront of the growing VR market. An estimated 90\% of Samsung Gear VR games and 53\% of Oculus Rift (games at launch) were Made With Unity. %Copy pasta from Unity about page.

    \subsubsection{Unreal vs Unity}
    % Why Unity over Unreal. Connect with IVR-Connection, UWP og MRT

    \subsubsection{Hololens}
    The HoloLens is a head-mounted display unit connected to an adjustable, cushioned inner headband, which can tilt HoloLens up and down, as well as forward and backward. To wear the unit, the user fits the HoloLens on their head, using an adjustment wheel at the back of the headband to secure it around the crown, supporting and distributing the weight of the unit equally for comfort, before tilting the visor towards the front of the eyes.

    In the front is much of the sensors and related hardware, including the cameras and processors. The visor is tinted; enclosed in the visor piece is a pair of transparent combiner lenses, in which the projected images are displayed in the lower half. %copy pasta wikipedia

    \subsubsection{Immersive Headsets}
    
